Parallel Information Set Generation for Kriegspiel

### 2012-06-25 sbac/paper.tex
Pared doc down to eight pages

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible game history is crucial to the
performance of game playing agents.  In this work, we discuss the problem of information set generation in the context
of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
Our results suggest that our algorithm can be efficiently parallelized.  
\end{abstract}

\Section{Introduction}
In imperfect information games, players do not have access to full knowledge of the world. Examples of imperfect
information include hidden cards in Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden tiles
in Scrabble~\cite{richards07opponent}, or hidden pieces in Stratego or Kriegspiel (partially observable
chess)~\cite{li94chess}. The game tree nodes that are indistinguishable to a player because they differ only in the
information that is hidden to the player by rule are called that player's {\em information set.}  The ability to
estimate the value of the possible states and to reason about the probability distribution over those states is crucial
to playing imperfect information games well.

@@

\Section{Background}
In this section, we describe the game of kriegspiel, and the particular variation of it that we have implemented.

Kriegspiel was invented by Henry Michael Temple in 1899.  Historically, the game has required three parties: two
competing players and one referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the
other.  One player controls the white pieces; the other controls the black pieces.  The referee has a board on which she
keeps track of both sets of pieces.  Each player has a full set of chess pieces.  He may position the opponent's pieces
however he wishes on his own board (if at all).  The players alternate moves as in a regular chess game, but instead of
announcing their moves out loud, they write the source and destination square of their desired plays on a piece of paper
and pass them to the referee, so that the requested move is not known to the other player. The referee, who knows the location of
all pieces, checks the legality of the move.  If the move is legal, the referee makes the move on his own board and
announces that a legal move has been made and that it is the other player's turn to move.  If the move is illegal
because it is blocked by an opponent's piece or would place or leave the active player in check, the referee announces,
``No.''  There is no penalty for attempting an illegal move.  The player continues to attempt moves until one finally
succeeds.  If a player has no legal moves (because of stalemate or checkmate), the referee announces the game's result.
With the advent of computers, it is convenient to play the game over a network, with a computer acting as the referee.

The referee makes other announcements besides declaring move illegal, and there are several variations of the game that
differ only in the nature of these announcements.  All declarations by the referee are heard by both players.  Thus, for
example, a player will hear and know if her opponent has attempted an illegal move, and will thus know that her
opponent's pieces are configured in such a way as to allow at least one attemptable move that blocked or would leave him
in check.

The referee announces when a player is in check.  (And again, this announcement is heard by both players.)  There is
some variation in what additional information is supplied in this case.  In one popular variant, the referee declares
that the player is in check ``by rank,'' ``by file,'' ``by diagonal,'' or ``by knight.''  In the case of a diagonal attack, the
referee may announce whether the checking piece is along the long or short diagonal (from the king's perspective).  It
is possible to be in check by two pieces at the same time, in which case the direction of both attacks would be
declared.  In our implementation, for simplicity, the referee declares only that the player is in check, without
specifying the nature of the attack.

In the case of the capture, the referee announces the location of the piece that was captured, so that both players can
remove the piece from the board.  In some variations, the referee also announces whether the captured piece was a pawn
or non-pawn.  In our implementation, information about the captured piece is not explicitly provided to the player who
makes the capture.

Because there is no penalty for making illegal moves, it is often to a player's advantage to attempt to make moves which
are likely to be illegal, as hearing such declarations from the referee can provide important information to a player
about the location of the opponent's pieces.  In particular, since a diagonal move by a pawn is legal only the case
where such a move would capture an opponent's piece, it could potentially be profitable to attempt all or many pawn
captures on every turn.  In order to speed the game up, popular variants of the game require the referee to make some
kind of declaration at the beginning of a player's turn with respect to the pawn captures that are available.  This may
be as simple as declaring ``Try'' to indicate that at least one pawn capture is possible.  In our implementation, the
referee declares all potential pawn captures.

In the case of a pawn promotion, the player making the pawn advancement would secretly notify the referee as to which
piece (queen, rook, bishop, or knight) he would like.  To reduce the number of bits needed to specify a move, our
implementation allows only promotion to queen.

In general, any move that is legal in a regular chess match may be at least attempted by a player in kriegspiel.  And
any kriegspiel move that is ultimately allowed by the referee would also be a legal move in chess.

We assume that a player may not attempt the same illegal move more than once per turn.  This means that if a player
hears multiple declarations from the referee that the opponent has attempted an illegal move, the player may assume that
the opponent has at least that many possibilities available to him.

\SubSection{Special Moves}
Our implementation supports the capability for pawns to move two spaces on their first turn.  However, we did not
implement support for castling or {\em en passant} pawn capture.  (These moves require maintaining additional bits of
information at each state that keep track of whether pieces have moved and/or whether pieces moved on the previous
turn.)

\SubSection{Example}
%\includegraphics[width=0.2\textwidth]{images/1W.png}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[1. d4]{\label{fig:1W}\includegraphics[width=0.2\textwidth]{images/1W.png}}
%\qquad
%\subfloat[1. ... a5]{\label{fig:1B}\includegraphics[width=0.2\textwidth]{images/1B.png}}
%\qquad
%\subfloat[2. Bg5]{\label{fig:2W}\includegraphics[width=0.2\textwidth]{images/2W.png}}
%\qquad
%\subfloat[2. ... b6]{\label{fig:2B}\includegraphics[width=0.2\textwidth]{images/2B.png}}
%\end{minipage}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[3. Nc3]{\label{fig:3W}\includegraphics[width=0.2\textwidth]{images/3W.png}}
%\qquad
%\subfloat[3. ... c5]{\label{fig:3B}\includegraphics[width=0.2\textwidth]{images/3B.png}}
%\qquad
%\subfloat[4. d5]{\label{fig:4W}\includegraphics[width=0.2\textwidth]{images/4W.png}}
%\qquad
%\subfloat[4. ... Na6]{\label{fig:4B}\includegraphics[width=0.2\textwidth]{images/4B.png}}
%\end{minipage}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[5. d6!]{\label{fig:5W}\includegraphics[width=0.2\textwidth]{images/5W.png}}
%\qquad
%\subfloat[5. ... f6!]{\label{fig:5B}\includegraphics[width=0.2\textwidth]{images/5B.png}}
%\qquad
%\subfloat[6. e4]{\label{fig:6W}\includegraphics[width=0.2\textwidth]{images/6W.png}}
%\qquad
%\subfloat[6. ... fg]{\label{fig:6B}\includegraphics[width=0.2\textwidth]{images/6B.png}}
%\end{minipage}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[7. Bc4]{\label{fig:7W}\includegraphics[width=0.2\textwidth]{images/7W.png}}
%\qquad
%\subfloat[7. ... ed]{\label{fig:7B}\includegraphics[width=0.2\textwidth]{images/7B.png}}
%\qquad
%\subfloat[8 Qd5!]{\label{fig:8W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
%\qquad
%\subfloat[8 ... Qc7?]{\label{fig:8B}\includegraphics[width=0.2\textwidth]{images/8B.png}}
%\end{minipage}
%\begin{minipage}{\textwidth}
%\centering
%\subfloat[9 Qf7+]{\label{fig:9W}\includegraphics[width=0.2\textwidth]{images/9W.png}}
%\qquad
%\subfloat[9 ... d8]{\label{fig:9B}\includegraphics[width=0.2\textwidth]{images/9B.png}}
%\qquad
%\subfloat[10 Qxf8++]{\label{fig:10W}\includegraphics[width=0.2\textwidth]{images/8W.png}}
%\end{minipage}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\end{figure}

@@

The concept of information sets is perhaps best explained through an example.  (See also Appendix A.)
Figures~\ref{listing} and~\ref{sequence} show a full listing and the associated pictorial progression of an actual
kriegspiel game from ~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves
are numbered in pairs: a move by white on one line followed by black's move on the next line.  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, black attempted three illegal moves.
\begin{figure}
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
\label{filteredlisting}
\end{figure}

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
the same sequence of observations for black is in black's information set.  Figure~\ref{abbrevoutput} shows the
information set for black after both player's have made five moves.  From this list of possibilities, black can infer
that white must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
black's observations up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for
each move in the game and the amount of time to find the full set on a single processor.  Appendix B shows the full
output of the program for the same point in the game (five moves for each player) and illustrates all of the possible
positions in the belief state.

@@

\begin{table}
\centering
\begin{tabular}{rrrrr}
 & \multicolumn{2}{c}{\bf White} & \multicolumn{2}{c}{\bf Black} \\
{\bf Ply} & {\bf Size} & {\bf Time} & {\bf Size} & {\bf Time} \\
1 & 18 & 0.000 & 20 & .000 \\
2 & 18 & 0.000 & 19 & .000 \\
3 & 18 & 0.000 & 404 & .008\\
4 & 279 & 0.008 & 401 & .024\\
5 & 242 & 0.028 & 1472 & .044\\
6 & 216 & 0.052 & 155 & .096\\
7 & 176 & 0.064 & 293 & .116\\
8 & 3406 & 0.092 & 158 & .128\\
9 & 2191 & 0.288 & 118 & .148\\
10 & 1423 & 0.508 & 25 & .152\\
11 & 1363 & 0.596 & 798 & .164\\
12 & 1416 & 0.744 & 518 & .192\\
13 & 1416 & 0.836 & 13564 & .304\\
14 & 3440 & 0.97 & 12394 & .776\\
15 & 3428 & 1.25 & 343652 & 3.75\\
16 & 50521 & 1.82 & 320704 & 17.3\\
17 & 43192 & 5.91 & 490162 & 99.9\\
18 & 26128 & 7.94 & 3792 & 119\\
19 & 19061 & 9.92 & 14836 & 121\\
\end{tabular}
\caption{Solution counts and running times for a sample kriegspiel game}
\label{bothtimes}
\end{table}

@@

\Section{The Information Set Generation Algorithm}
We have formulated information set generation as a combinatorial search problem.  The input to the algorithm is a list
of observations for one player in a kriegspiel game.  We assume that the agent has perfect recall (i.e., remembers all
of its own decisions), which is a reasonable assumption for kriegspiel.  (See~\cite{gilpin07algorithms} for a further
discussion of perfect recall.)

The nodes in the search tree at depth $n$ correspond possible to possible sequences of $n$ plies from the start state.
There is a search space variable for each ply in the game, and the values that a variable can take on are all possible
legal moves.  A kriegspiel game may last for on the order of tens of plies.  Moves are parameterized by the piece that
is relocating (both black and white have six different types of pieces: king, queen, rook, bishop, knight, pawn), the
source square, and the destination square.  Only a few hundred of the $64 \times 64 \times 6$ combinations represent
moves that are legitimate in at least some scenarios.  For example, no piece could ever move from a3 to h2\footnote{We
observe the convention that on an $8 \times 8$ chessboard the rows are numbered 1 to 8 and the columns are labeled a to
h.} in one move.

Information set generation, then, is an all-solutions search through the space.  Whenever a node in the search tree is
found to be inconsistent with the actual observation given in the input, the subtree rooted at that node is pruned.
There are types of observations that must match in order for the state to be found: check status, pawn tries, and 
illegal moves available/attempted.

Figure~\ref{codelisting} shows the function at the heart of our information set generation algorithm.  The actual
observations can be viewed as global, read-only variables.  All solutions (i.e., all nodes in the information set) will
be located at the same depth in the tree: the depth that is equal to the number of plies in the observation list.
Referee announcements regarding pawn tries or a player being in check are made prior to a move.  The algorithm checks
for the consistency of these observations in lines $5--6$.  If the search reaches the appropriate depth and passes these
tests, then a solution has been found and is reported (line $7$).

Otherwise, we consider the possible moves that can be made from that state.  If the player whose turn it is to move at
the current level of the search tree is the same player from whose perspective we have received observations, then the
next step is to make sure that every legal move that was attempted at the player at that ply is also at least
attemptable at the game position that corresponds to the game tree node (lines $15--22$).  We know exactly the legal
move that was ultimately made at this depth (because we are assuming perfect recall), and we check to make sure that
this move is indeed legal (pruning otherwise).  If so, we apply that move and recurse (lines $24--32$).  This
corresponds to a ``forced move'' in the general search paradigm.

On the other hand, if the player whose turn it is to move at the current depth of the search tree is the opponent, then
must consider all possible legal moves to branch on.  We first check to make sure that the number of legal moves from
this position is at least as large as the number of illegal moves that the opponent attempted (i.e., the number of times
the referee said ``No'' before finally accepting the opponent's move).  This is shown in lines $39--42$.  At this point
we generate a child for each move that would be accepted by the referee and recurse (lines $44--52$).

One of our goals was to assess the varying utilities of the different pruning criteria.  We tried some different
orderings of the tests and were somewhat surprised (and disappointed) to see that there was actually a fairly minimal
impact on performance.  This could be due to the fact that because of the tedious nature of encoding problem instances
from actual games, we were not able to run experiments of a wide variety of types of positions (i.e., different styles
of play).

@@

\Section{Charm++}
We built our search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms.

@@

\Section{The {\sc Charm++} Parallel State Space Search Engine}
Sun \etal have developed the Parallel State Space Search Engine ({\sc ParSSSE})
within the {\sc Charm++} framework, which exposes an API that is useful for a
variety of graph-based and tree-based search problems~\cite{sun11adaptive}. In
{\sc ParSSE} the fundamental unit of work is the processing of a node.
Generally, this means identifying whether the node corresponds to a solution
and/or generating descendant nodes and spawning work tasks related to them. A
task is {\em terminal} if it spawns no children.

@@

\Section{Performance Results}
We built our search on top of the same general search engine that we used in class.  The performance results reported in
this section were obtained by running our program on Bleuprint with up to 1024 processors.

Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
the size of the information set was about 4.5 million.  (This instance is not necessarily representative of all games or
states of the game, but it was not cherry-picked in any way.)  The table shows that the problem is indeed highly
parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.

\begin{table}
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}
\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
\label{speedups}
\end{table}

\begin{figure}[h]
\centering
\includegraphics[width=0.7\columnwidth]{plots/3schemes.pdf}
\caption{Experiments using different load balancing strategies.}
\label{3schemes}
\end{figure}

In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
variation in performance across the different strategies.  We did notice that across all of our experiments, the work
stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
redistribution.  This is probably also related to the fact that we are doing an all-solutions search.

\begin{figure}[h]
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\caption{(Left) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Right) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
\label{3depths}
\end{figure}

>>>>

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible game history is critical to the
performance of game playing agents.  In this work, we address the parallelization of information set generation in the context
of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
We achieve speedups of over 500 on 1024 processors, far exceeding previous scalability results for game tree search applications.
\end{abstract}

\Section{Introduction}
In imperfect information games, players do not have access to full knowledge of
the world. Examples of imperfect information include hidden cards in
Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
tiles in Scrabble~\cite{richards07opponent}, or hidden pieces 
Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
that are indistinguishable to a player because they differ only in the
information that is hidden to the player by rule compose that player's {\em
information set.}  The ability to estimate the value of the possible states and
to reason about the probability distribution over those states is crucial to
playing imperfect information games well~\cite{thielscher12hyperplay}.

Information Set Generation (ISG) is an operation that game tree search agents
can use to identify nodes in the game tree that are consistent with their
observations~\cite{richards12reasoning}.  Because the operation is relatively
expensive and may be repeatedly applied by a decision-making agent, its
efficiency is a key factor in an agent's overall performance.  In this work, we
show that ISG for kriegspiel can be efficiently parallelized.

@@

The paper is organized as follows.  In Section~\ref{Background}, we review necessary background information, including the rules of kriegspiel and the nature of the information set generation problem.  In Section ~\ref{ParSSSE}, we describe the {\sc Charm++} Parallel State Space Search Engine (ParSSSE) platform.  We then provide our experimental results in Section~\ref{Results}.  Related work is discussed in Section~\ref{Related}.  We conclude with a discussion of future work.

\Section{Background}\label{Background}
\subsection{Kriegspiel}
Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
game has required three parties: two competing players and one referee.  Today,
players often compete over a computer network, with the role of the referee
being automated.  One player sees and controls only the white pieces; the
other, black.  The referee track of both sets of pieces.  The players alternate
moves as in a regular chess game, but instead of openly declaring their moves,
they secretly communicate move requests to the referee.  The referee, who knows
the location of all pieces, checks the legality of the move.

If the move is legal, the referee makes executes the move and announces that a
legal move has been made and that it is the other player's turn to move.  If
the move is illegal because it is blocked by an opponent's piece or would place
or leave the active player in check, the referee announces, ``No.''  The player
then continues, without penalty, to attempt moves until one finally succeeds.
If a player has no legal moves because of stalemate or checkmate, the referee
announces the game's result.

The referee makes other announcements besides declaring move illegal, and there
are several variations of the game that differ only in the nature of these
announcements.  All declarations by the referee are made to both players.  In
particular, a player will hear and know if her opponent has attempted an
illegal move, and will thus know that her opponent's pieces are configured in
such a way as to allow at least one attemptable move that is blocked or would
leave him in check.

The referee announces when a player is in check. In the case of the capture,
the referee announces the location of the piece that was captured but not its
identity.  Because there is no penalty for making illegal moves, it is often to
a player's advantage to attempt to make moves which are likely to be illegal,
as hearing such declarations from the referee can provide important information
to a player about the location of the opponent's pieces.  In particular, since
a diagonal move by a pawn is legal only the case where such a move would
capture an opponent's piece, it could potentially be profitable to attempt all
or many pawn captures on every turn.  In order to speed the game up, the
referee declares, at the beginning of a player's turn, all locations where a
pawn capture is possible.

In general, any move that is legal in a regular chess match may be at least
attempted by a player in Kriegspiel.  And any Kriegspiel move that is
ultimately allowed by the referee would also be a legal move in chess.

\SubSection{Example}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.5\linewidth}
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\end{figure}

@@

\subsection{Information Set Generation}
The concept of information sets is perhaps best explained through an example. 
Figure~\ref{listing} and show a full listing and the associated pictorial progression of an actual
kriegspiel game~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, black attempted three illegal moves.
\begin{figure}
\begin{minipage}[b]{0.35\linewidth}
\centering
\small
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
\end{verbatim}
\end{minipage}
\begin{minipage}[b]{0.25\linewidth}
\small
\centering
\begin{verbatim}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\end{minipage}
\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
\label{filteredlisting}
\end{figure}

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
the same sequence of observations for black is in black's information set.  Figure~\ref{abbrevoutput} shows the
information set for black after both player's have made five moves.  From this list of possibilities, black can infer
that white must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
black's observations up to that point.

%Table~\ref{bothtimes} shows the size of the information sets for each player for
%each move in the game and the amount of time to find the full set on a single processor.  Appendix B shows the full
%output of the program for the same point in the game (five moves for each player) and illustrates all of the possible
%positions in the belief state.

@@

%\begin{table}
%\centering
%\begin{tabular}{rrrrr}
% & \multicolumn{2}{c}{\bf White} & \multicolumn{2}{c}{\bf Black} \\
%{\bf Ply} & {\bf Size} & {\bf Time} & {\bf Size} & {\bf Time} \\
%1 & 18 & 0.000 & 20 & .000 \\
%2 & 18 & 0.000 & 19 & .000 \\
%3 & 18 & 0.000 & 404 & .008\\
%4 & 279 & 0.008 & 401 & .024\\
%5 & 242 & 0.028 & 1472 & .044\\
%6 & 216 & 0.052 & 155 & .096\\
%7 & 176 & 0.064 & 293 & .116\\
%8 & 3406 & 0.092 & 158 & .128\\
%9 & 2191 & 0.288 & 118 & .148\\
%10 & 1423 & 0.508 & 25 & .152\\
%11 & 1363 & 0.596 & 798 & .164\\
%12 & 1416 & 0.744 & 518 & .192\\
%13 & 1416 & 0.836 & 13564 & .304\\
%14 & 3440 & 0.97 & 12394 & .776\\
%15 & 3428 & 1.25 & 343652 & 3.75\\
%16 & 50521 & 1.82 & 320704 & 17.3\\
%17 & 43192 & 5.91 & 490162 & 99.9\\
%18 & 26128 & 7.94 & 3792 & 119\\
%19 & 19061 & 9.92 & 14836 & 121\\
%\end{tabular}
%\caption{Solution counts and running times for a sample kriegspiel game}
%\label{bothtimes}
%\end{table}

@@

{The Information Set Generation Algorithm}
ISG can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of observations for one player in a kriegspiel game.

The nodes in the search tree at depth $n$ correspond possible to possible sequences of $n$ plies from the start state.
There is a search space variable for each ply in the game, and the values that a variable can take on are all possible
legal moves.  Information set generation, then, is an all-solutions search through the space.  Whenever a node in the search tree is
found to be inconsistent with the actual observation given in the input, the
subtree rooted at that node is pruned (i.e., search is discontinued along that
path).  There are types of observations that must match in order for the state
to be found: check status, pawn tries, and illegal moves available/attempted.

Figure~\ref{codelisting} shows the function at the heart of our information set generation algorithm.  The actual
observations can be viewed as global, read-only variables.  All solutions (i.e., all nodes in the information set) will
be located at the same depth in the tree: the depth that is equal to the number of plies in the observation list.
Referee announcements regarding pawn tries or a player being in check are made prior to a move.  The algorithm checks
for the consistency of these observations at line $7$.  If the search reaches the appropriate depth and passes these
tests, then a solution has been found and is added to the information set (line $3$).

%Otherwise, we consider the possible moves that can be made from that state.  If the player whose turn it is to move at
%the current level of the search tree is the same player from whose perspective we have received observations, then the
%next step is to make sure that every legal move that was attempted at the player at that ply is also at least
%attemptable at the game position that corresponds to the game tree node (lines $15--22$).  We know exactly the legal
%move that was ultimately made at this depth (because we are assuming perfect recall), and we check to make sure that
%this move is indeed legal (pruning otherwise).  If so, we apply that move and recurse (lines $24--32$).  This
%corresponds to a ``forced move'' in the general search paradigm.
%
%On the other hand, if the player whose turn it is to move at the current depth of the search tree is the opponent, then
%must consider all possible legal moves to branch on.  We first check to make sure that the number of legal moves from
%this position is at least as large as the number of illegal moves that the opponent attempted (i.e., the number of times
%the referee said ``No'' before finally accepting the opponent's move).  This is shown in lines $39--42$.  At this point
%we generate a child for each move that would be accepted by the referee and recurse (lines $44--52$).
%
%One of our goals was to assess the varying utilities of the different pruning criteria.  We tried some different
%orderings of the tests and were somewhat surprised (and disappointed) to see that there was actually a fairly minimal
%impact on performance.  This could be due to the fact that because of the tedious nature of encoding problem instances
%from actual games, we were not able to run experiments of a wide variety of types of positions (i.e., different styles
%of play).

@@

\Section{Charm++ and ParSSSE}\label{ParSSSE}
We built our search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms.

@@

%\subsection{The {\sc Charm++} Parallel State Space Search Engine}
Sun \etal have developed the Parallel State Space Search Engine ({\sc ParSSSE})
within the {\sc Charm++} framework, which exposes an API that is useful for a
variety of graph-based and tree-based search problems~\cite{sun11adaptive}. In
{\sc ParSSE} the fundamental unit of work is the processing of a node.
Generally, this means identifying whether the node corresponds to a solution
and/or generating descendant nodes and spawning work tasks related to them. A
task is {\em terminal} if it spawns no children.

@@

%%We built our search on top of the same general search engine that we used in class.  The performance results reported in
%%this section were obtained by running our program on Bleuprint with up to 1024 processors.
%%
%%Table~\ref{speedups} shows the performance results for our algorithm on a randomly selected problem instance in which
%%the size of the information set was about 4.5 million.  (This instance is not necessarily representative of all games or
%%states of the game, but it was not cherry-picked in any way.)  The table shows that the problem is indeed highly
%%parallizable.  The speedups are nearly linear for up  to 32 processors, and even on 1024 processors, we get a good
%%speedup.  In this particular instance, the sequential algorithm takes almost 25 minutes to run.  But when parallelized
%%on 1024 processors---perfectly reasonable resource usage on modern hardware---the search takes less than three seconds.
%%This is a speedup of 576 times!  It might not be reasonable to use the algorithm at all if it takes 25 or more to run
%%for some turns.  But if the cost is only a few seconds, it becomes entirely feasible.
%%
%%\begin{table}
%%\centering
%%\begin{tabular}{rrr}
%%{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
%%1	&	1524	&	1\\
%%2	&	812	&	1.88\\
%%4	&	407	&	3.74\\
%%8	&	208	&	7.32\\
%%16	&	102	&	14.8\\
%%32	&	53.2	&	28.7\\
%%64	&	28.0	&	54.5\\
%%128	&	14.9	&	102\\
%%256	&	7.70	&	197\\
%%512	&	4.59	&	331\\
%%1024	&	2.64	&	576\\
%%\end{tabular}
%%\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
%%10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
%%\label{speedups}
%%\end{table}
%%
%%\begin{figure}[h]
%%\centering
%%\includegraphics[width=0.7\columnwidth]{plots/3schemes.pdf}
%%\caption{Experiments using different load balancing strategies.}
%%\label{3schemes}
%%
%%In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
%%used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
%%engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
%%variation in performance across the different strategies.  We did notice that across all of our experiments, the work
%%stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
%%that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
%%the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
%%redistribution.  This is probably also related to the fact that we are doing an all-solutions search. 
%%
%% 
%%\begin{figure}[h]
%%\centering
%%\begin{minipage}{0.5\linewidth}
%%\includegraphics[width=\columnwidth]{plots/3depths.pdf}
%%\end{minipage}
%%\centering
%%\begin{minipage}{0.5\linewidth}
%%\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
%%\end{minipage}
%%\caption{(Left) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
%%(Right) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
%%\label{3depths}
%%
%%We next attempted to measure the isoefficiency of the problem---how is efficiency of the parallelization of the
